\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{fullpage}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}

\geometry{top=1in, bottom=1in, left=1in, right=1in}

\DeclareMathOperator{\sgn}{sgn}

\title{CS 393R Autonomous Robots \\ \large Assignment 1: Obstacle Avoidance}
\author{Elvin Yang, Aidan Dunlap}
\date{September 20, 2021}

\begin{document}
\maketitle

\section{Questions}

Let
\begin{itemize}
    \item
    $l$, $w$, $b$, $d$, and $m$ represent the car's length, width, wheel base,
    track width, and obstacle safety margin, respectively.

    \item
    $r$ represent the car's radius of turning. Negative $r$ signifies a right
    turn.

    \item
    $p = (x, y)$ represent an observed point in the base\_link reference frame.

    \item
    $a = \frac{l - b}{2}$ represent the distance between the front axle and the
    front end of the car as well as the distance between the rear axle and the
    rear end of the car.

    % \item
    % $c_1, c_2, c_3, c_4$ represent the corners of the car in each respective
    % quadrant of the base\_link reference frame.
\end{itemize}

\begin{enumerate}[leftmargin=*]
    \item
    Which point of the car traces an arc of maximum radius? What is that radius,
    in terms of $r$ and other car parameters?

    The front corner of the car that is farthest from the center of turning
    traces an arc of maximum radius. It has radius
    $$ r_{max} = \sgn(r) \sqrt{\left(l + a\right)^2 + \left(\lvert r \rvert + \frac{w}{2}\right)^2} $$

    \item
    Which point of the car traces an arc of minimum radius? What is that radius,
    in terms of $r$ and other car parameters?

    The rear corner of the car closest to the center of turning traces an
    arc of minimum radius. It has radius
    $$ r_{min} = \sgn(r) \sqrt{\left(l - a \right)^2 + \left(\lvert r \rvert - \frac{w}{2}\right)^2} $$

    \item
    Under what conditions will $p$ first hit the inner side (the side closer to
    the center of turning) of the car as the car drives forward?

    Let $r_1, r_2$ represent the distances from the endpoints of the inner side
    of the car to the center of turning, where $r_1 < r_2$.  Let $d_p$ represent
    the distance from $p$ to the center of turning.

    If $r_1 < d_p < r_2$, then $p$ will hit the inner side of the car.

    \item
    Under what conditions will $p$ first hit the front of the car as the car
    drives forward?

    If the car is driving straight, let $y_{min}, y_{max}$ represent the
    y-values of the front corners of the car in the base link frame. $p$ will
    hit the car if $y_{min} < p_y < y_{max}$.

    If the car is making a turn, let $r_1, r_2$ represent the distances from the
    endpoints of the front side of the car to the center of turning, where
    $r_1 < r_2$.  Let $d_p$ represent the distance from $p$ to the center of
    turning. If $r_1 < d_p < r_2$, then $p$ will hit the front side of the car.

    \item
    Under what conditions will $p$ first hit the outer side (the side farther to
    the center of turning) of the car as the car drives forward?

    The instantaneous direction of a point on the car is perpendicular to the
    vector from the center of turning to that point. Without loss of generality,
    for left turns, the instantaneous velocity of the rear corner of the outer
    side has a negative $y$ component and the instantaneous velocity of the
    front corner of the outer side has a positive $y$ component.

    If $p$ is sufficiently close to the rear corner of the outer side of the
    car, it will hit the outer side of the car. The area of the region in which
    a point will hit the outer side of the car increases as the difference in
    radii between the base link and outer rear corner increase.

    % mathematical explanation as well?
    % y in -a to a
    % x max is r_3 - |r|

    \item
    What is the maximum distance (the free path length) the car can move forward
    along the arc before it hits the point $p$?

    Find the angle between the hit point and the vector from the base of the car to
    the turning point, called $c$ (math in code). Use the method discussed in
    questions 3-4 to determine if a point will hit the car at all. If so, handle the
    next step by cases depending on whether there was a hit on the side or front of
    the car.

    If there was a hit on the front of the car, approximate the hit point on the
    car as follows: let $p = \frac{d_1 - r_1}{r_2 - r_2}$, where $d_1$ is the
    distance between the collision point and the turning point. Let $c_1$ be the
    south-east of the car, and $c_2$ the north-east. We can calculate $\beta$,
    the angle between the base link and the hit point projected onto the car,
    calculated as $c_1 + p * \vec{c_1 c_2}$  in the base link, using the same
    method as per $c$. Using this, calculate the angle from the base frame to
    the base frame on collision, $\alpha = c - \beta$.     For side hits, do a
    similar process, or approximate a standard hit point like the middle of the
    side given the smaller area for points to hit at a time. For negative radii,
    do about the same thing except for nuances in vector angle calculations
    (also in code).

    The free path length is $\alpha \lvert r \rvert$, the arc length of the
    angle subtended by the base link.

    \item
    If the current velocity of the car is $v$, and the maximum magnitude of
    deceleration is $a$, what is the minimum stopping distance of the car?
    \begin{align*}
        \lVert v_f \rVert^2 &= \lVert v_i \rVert^2 - 2ad_{stop} \\
        0 &= \lVert v \rVert^2 - 2 a d_{stop} \\
        d_{stop} &= \frac{\lVert v \rVert^2}{2a}
    \end{align*}
\end{enumerate}

\section{Algorithm Details}

\subsection{Parameters}

We define the following as constants:
\begin{itemize}
    \item Size of the car.
    \\ We measured these with a tape measure. We used these for a variety of
    uses from collision detection to path planning, multiplied by a
    margin-of-safety coefficient (c > 1) that increased the size of the car.
    \item Safety margin around the car.
    \\ We manually searched for a safety margin  coefficient that we thought
    managed both safety and flexibility effectively.
    \item Limits of the car's steering angle.
    \\ We measured this with a tape measure with the car's maximum controllable
    turning radius. This was used for a lower/upper bound on the car's turning
    angle which we used to search for potential paths.
    \item Steering angle step size for testing different constant-curvature arcs.
    \\ This was a hand-tuned parameter that we adjusted to balance turning
    flexibility and route calculation latency
    \item The maximum magnitude of velocity, acceleration, and deceleration.
    \\ We used this to plan our time-optimal control paths. We never used or
    tested the maximum velocity of the car and instead stuck to safe but also
    challenging velocities.
    \item Actuation latency.
    \\ We used this for latency compensation (see more below)
    \item Location of the LIDAR sensor.
    \\ We used this to translate LIDAR points back into the base link/predicted
    base link reference frame.
\end{itemize}

Our path-selection algorithm uses the following properties of each path option:
\begin{itemize}
    \item The free path length.
    \\ The distance the car could travel such that it both does not collide with
    an obstacle and also doesn't drive past the closest point on the circle to
    the target.
    \item The point on the free path closest to the navigation goal.
    \\ We used this to determine the smallest absolute displacement between the
    car on its path and the target.
\end{itemize}

\subsubsection{Parameter Tuning}

We measured the physical dimensions of the car and the location of the LIDAR
sensor in the base link reference frame. We tuned actuation latency by trial and
error, testing multiple latencies until the car was able to travel in a straight
path of 2 meters within an acceptable margin of error.

\subsection{Limitations and Improvements}

We observed that our latency compensation was able to achieve a relatively low
displacement error when we only take into account actuation latency. However, we
noticed that the navigation target would shift a lot in the visualization. We
are not sure whether this is due to observation latency or noise and precision
errors in our transformation code, and we plan on correcting for observation
latency to rule it out as a factor.

\bigskip
\noindent
We observed that our path-finding algorithm worked well without factoring in
clearance to the path-selection heuristic since the safety margin enforces an
lower bound on the clearance for each path. However, we plan on merging
clearance code into our main development branch to make our algorithm more
robust in complex environments.

% code performance?

\subsection{Challenges}

Upon running the car for the first time, we found that the steering was heavily
biased to the left. We had to set the
\texttt{steering\_angle\_to\_servo\_offset} higher than the suggested range for
steering to be symmetric.

We encountered a lot of bugs due to infinite radii in the straight-line-path
case. Most of our functions perform separate computations for this edge case.

\section{Repository Link}
\href{https://github.com/elvout/cs393r-a1}{https://github.com/elvout/cs393r-a1}

\section{Demonstration}
Three videos demonstrating obstacle avoidance can be found
\href{https://drive.google.com/drive/folders/1Z8qGX_YRjMb96KEpW027cGvmGeLOB8BU}{here}.

\noindent
In all scenarios, the car's navigation target is directly in front of it. In the
first video, the car must weave between bins to reach its target. The second
video shows the car taking a wide arc to avoid three obstacles lined up directly
in front of the starting position. In the third scenario, the car chooses an
opening in a wall obstacles to travel through.

\section{Contributions}

When we began, Aidan worked on obstacle detection and path selection and Elvin
worked on time optimal control and latency compensation. We spent a significant
amount of time working as a group to figure out math, integrate our
components together, and test our code on the physical car.

\end{document}
